---
title: "Bidirected Graphs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Bidirected_Graphs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This is replication code for the manuscript by Hu and Evans (2022).
```{r setup}
library(imset)
```

## Number of Bidirected Graphs that do not define the model

```{r bidi5, eval=TRUE}
data(bidi5)

dm <- logical(length(bidi5))
for (i in seq_along(bidi5)) {
  dm[i] <- defines_mod(bidi5[[i]])
}

cat(sum(dm), " imsets define the model out of ", length(dm), "\n", sep="")
```
Which graph fails to define the model?
```{r 5cycle}
bidi5[[which(!dm)]]
```

## Bidirected Graphs of size 6

We can take the same approach for graphs of size 6, but we don't evaluate it
here for reasons of time.

```{r bidi6, eval=FALSE}
data(bidi6)

dm <- logical(length(bidi6))
for (i in seq_along(bidi6)) {
  dm[i] <- defines_mod(bidi6[[i]])
}

cat(sum(dm), " imsets define the model out of ", length(dm), "\n", sep="")
```

## Testing Decomposition

We can also check that the decomposition given in Section 5 gives the same 
results as the integer linear program.

```{r, eval=TRUE}
dmZH <- logical(length(bidi5))

for (i in seq_along(bidi5)) {
  dmZH[i] <- test_decomp_ZH(bidi5[[i]])
}

which(dm != dmZH)
```

