---
title: "Introdution"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introdution}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Imsets

This package allows for the construction, manipulation and testing of independence
models represented by _imsets_.  

An _imset_ is an **i**nteger-valued **m**ulti**set**, or a function from the 
powerset of a collection of variables $V$ to the integers.  They can be used
to represent arbitrary conditional independence models, by assessing whether or
not certain imsets are _structural_.  

### Basic Definitions

The _identifier_ imset $\delta_A$ is a vector indexed by subsets of $V$ with 
a 1 in the entry that corresponds to $A$, and 0 in all other entries.

A _semi-elementary imset_ is isomorphic to a triple $A,B,C$ of disjoint sets 
where the order of the first two elements is arbitrary.  It is defined as
\begin{align*}
u_{\langle A,B | C\rangle} := \delta_{C} - \delta_{AC} - \delta_{BC} + \delta_{ABC}
\end{align*}

An imset $u$ is said to be _structural_ if there is some natural number $k \in \mathbb{N}$
such that $k\cdot u$ can be written as a sum of semi-elementary imsets with 
natural numbers as coefficients.  In other words, if 
\begin{align}
\exists k \in \mathbb{N} : \qquad k \cdot u = \sum_{i} k_i \cdot u_{\langle A_i,B_i | C_i\rangle} \qquad \text{for } k_i \in \mathbb{N}. \label{eqn:struc}
\end{align}
An _imset_ is said to _represent_ an independence $A \perp B \mid C$ if there 
exists a $k \in \mathbb{N}$ such that
\begin{align*}
k \cdot u - u_{\langle A,B | C\rangle}
\end{align*}
is structural.  Note that therefore the imset itself must be structural, and will 
represent any of the independences in the expansion on the right-hand side of 
(\ref{eqn:struc}). 

### The Package

We can load the package in the usual manner.

```{r setup}
library(imset)
```

Then to define an imset, we can use the `elem_imset` function.
```{r elem_imset}
elem_imset(1,2)
elem_imset(1,2,4)
elem_imset(1,2,4,n=6)
```
Note that the argument `n` modifies the total number of variables, which is 
otherwise assumed to be the largest number specified.

We can test whether a given imset is structural using `is_structural`.
```{r is_structural}
up <- elem_imset(1,2) + elem_imset(1,3,2)
um <- elem_imset(1,2) - elem_imset(1,3,2)
is_structural(up)
is_structural(um)
```
Note that if the imset is structural, then a valid decomposition is returned.  It 
is clear that this is not unique in general.

### Standard Imsets

The standard imset for a DAG is defined in the usual way.  We can create a 
DAG using the `graphCr` function from the `MixedGraphs` package.
```{r makeDAG}
dag1 <- graphCr("1 -> 3 <- 2, 3 -> 4")
u1 <- standard_imset(dag1)
u1
```
The characteristic imset can be obtained by applying `char_imset` to either the 
graph or the standard imset.

We can also check whether the standard imset defines the model for the DAG.
```{r defines_mod}
defines_mod(dag1, u1)
```
And then test another imset which we know does not:
```{r ndefines_mod}
defines_mod(dag1, elem_imset(1,4))
```
